Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-08-14 manuscript/main.tex
compact proof easily done

	\subsection{Tucker's Criterion}\jongmin{notation not yet changed}
	We derive that the PLS components obtained from the first step of the proposed algorithm satisfy the Tucker's Criterion \citep{Tucker1938} extended to our scalar-on-hybrid regression model setting \eqref{eq: Hybrid functional model}. We omit $l$ in the notations and first define the cross-covariance term between the response and the functional predictors (observed versions if $l=1$ and residualized versions if $l \ge 2$): 
	\begin{equation}\label{def:cross_covariance_yx}
\sigma_{YX}
=
 \bigl(
 \sigma_{YX, 1}(t) , \ldots, \sigma_{YX, K}(t) 
 \bigr)
  :=
\bigl(
\mathbb{E}
[Y_1 X_{11}]
,\ldots, 
\mathbb{E}
[Y_1  X_{1K}]
\bigr)   \in \mathcal{F}^{\otimes K},
	\end{equation}
and between the response and the scalar predictor:
\begin{equation*}
\sigma_{YZ} 
= 
(\sigma_{YZ,1}, \ldots, \sigma_{YZ,p})^\top 
:= 
\bigl(
\mathbb{E}[Y_1 Z_{11}]
,\ldots,
\mathbb{E}[Y_1 Z_{1p}]
\bigr)^\top %= E(Y\mathbf{Z})
\in \mathbb{R}^p.
\end{equation*}
Based on these definitions, we define the cross-covariance term between the response and the hybrid predictor as
\begin{equation*}
\Sigma_{YW} :=  
\mathbb{E}
[ Y_1 W_1 ]
=
\bigl(
\sigma_{YX, 1} , \ldots, \sigma_{YX, K} , \, \sigma_{YZ}
\bigr) \in \mathcal{H}.
\end{equation*}
Based on these definitions, we introduce two cross-covariance operators.
\begin{lemma}\label{lemma:cross_cov_functional}
Define the operator $
\mathcal{C}_{YW}
=
\mathbb{E}(W_1 \otimes_\mathcal{H} Y_1)
: \mathcal{H} \rightarrow \mathbb{R}$ such that for any $h=(f_1, \ldots, f_K, \mathbf{v}) \in \mathcal{H}$, it maps $h$ to a real number as: \begin{equation*} \mathcal{C}_{YW} h := \mathbb{E} \left[ \langle W_1, h \rangle_\mathcal{H} Y_1 \right] = \langle \Sigma_{YW}, h \rangle_\mathcal{H} = \sum \limits_{k=1}^K \int_0^1 \sigma_{YX, k}(t) \, f_k(t) \, dt + \sigma_{YW}^\top \mathbf{v}. \end{equation*} 
If there exist finite constants $Q_1$ and $Q_2$ such that
\begin{align}\label{condition_for_compact}
	\max \limits_{k=1, \ldots, K} \sup \limits_{t \in [0,1]} \sigma_{YX, k}^{2}(t) < Q_1 \quad \textnormal{and} \quad  \max_{r=1, \ldots, p} \sigma_{YZ, r}^2 < Q_2,
\end{align}
the operator $\mathcal{C}_{YW}$ is a compact operator.
\end{lemma}
Proof of Lemma \ref{lemma:cross_cov_functional}
is provided in Appendix \ref{section:proof:lemma:cross_cov_functional}.
\begin{lemma}\label{lemma:cross_cov_operator}
Define $\mathcal{C}_{YW}= : \mathcal{H} \rightarrow \mathbb{R}$,
which	maps  
	$h=(f_1, \ldots, f_K, \mathbf{v}) \in \mathcal{H}$ to a real number:
	\begin{equation*}
		\mathcal{C}_{YW}
		h
		:=
		\mathbb{E} 
		\bigl[ 
		\langle W_1, h \rangle_\mathcal{H} Y_1 
		\bigr] 
		= 
		\langle \Sigma_{YW}, h \rangle_\mathcal{H}
		=
		\sum \limits_{k=1}^K \int_0^1 \sigma_{YX, k}(t) \, f_k(t) \, dt 
		+ 
		\sigma_{YW}^\top \mathbf{v}.
	\end{equation*}
	Next, define $\mathcal{C}_{WY} = \mathbb{E}[Y \otimes W]: \mathbb{R} \rightarrow \mathcal{H}$,
which	maps  $d \in \mathbb{R}$ to a hybrid object in $\mathcal{H}$ as follows:
	%%%%
	\begin{align*}		
		\mathcal{C}_{WY} d 
		:= 
		\mathbb{E} 
		\bigl[
		\langle Y_1, d \rangle W_1 
		\bigr] 
		= 
		\mathbb{E}
		[ Y_1 W_1 d ] =  d \, \Sigma_{Y W}.
	\end{align*}
Then	$\mathcal{C}_{WY}$ is an adjoint operator of $\mathcal{C}_{YW}$. That is, $\mathcal{C}_{WY} = \mathcal{C}_{YW}^\ast$.
\end{lemma}

The proof of Lemma \ref{lemma:cross_cov_operator} is provided in Appendix \ref{section:proof:lemma:cross_cov_operator}.
Based on these two operators, we define a positive   operator as follows:
\begin{lemma}[Composite cross-covariance operator]\label{lemma:composite_cross_cov}
Define
	$\mathcal{U} := \mathcal{C}_{WY} \circ \mathcal{C}_{YW}: \mathcal{H} \rightarrow \mathcal{H}$ as an operator which performs the following mapping:
	\begin{equation*}
		\mathcal{U} h 
		= 
		\mathcal{C}_{WY} ( \mathcal{C}_{YW} h) 
		=
		\mathcal{C}_{YW} (\langle \Sigma_{YW}, h \rangle_\mathcal{H}) 
		=  
		\Sigma_{YW} \, \langle \Sigma_{YW}, h \rangle_\mathcal{H} 
		= 
		(\Sigma_{YW} \otimes \Sigma_{YW}) h.
	\end{equation*}
		In other words, $\mathcal{U} = \Sigma_{YW} \otimes \Sigma_{YW}$.
		Then $\mathcal{U}$
	 is self-adjoint and positive-semidefinite.
\end{lemma}
The proof of Lemma \ref{lemma:composite_cross_cov} is provided in Appendix \ref{section:proof:lemma:composite_cross_cov}.

@@

\section{Proof of Lemma \ref{lemma:composite_cross_cov}}\label{section:proof:lemma:composite_cross_cov}
\begin{proof}
	The operator
$\mathcal{U}$ is \textit{self-adjoint} because for any $h_1, h_2 \in \mathcal{H}$, we have
\begin{align*}
\langle 
\mathcal{U} h_1, \, h_2 
\rangle_\mathcal{H}  
&= 
\bigl \langle 
	\langle h_1, \Sigma_{YW} \rangle_\mathcal{H} \Sigma_{YW},
	\,
	h_2
\bigr \rangle_\mathcal{H}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\\&= \langle h_1, \Sigma_{YW} \rangle_\mathcal{H} \langle \Sigma_{YW}, h_2 \rangle_\mathcal{H}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\\&= 
\bigl \langle 
h_1,
\,
 \langle \Sigma_{YW}, h_2 \rangle_\mathcal{H} \Sigma_{YW} 
\bigr \rangle_\mathcal{H}
\\ 	&= \langle h_1, \mathcal{U} h_2 \rangle_\mathcal{H}.
\end{align*}
Next, 
	$\mathcal{U}$ is \textit{positive-semidefinite} because for every $h \in \mathcal{H}$, we have
\begin{equation*}
\langle 
	\mathcal{U} h,
	\,
	h
\rangle_\mathcal{H}
=
\bigl \langle 
	\langle 
		\Sigma_{YW},\, h \rangle_{\mathcal{H}}  \Sigma_{YW}, \, h
\bigr \rangle_\mathcal{H}
	= \langle \Sigma_{YW}, \, h \rangle_\mathcal{H}^2 \ge 0.
\end{equation*}
This completes the proof of Lemma \ref{lemma:composite_cross_cov}.
\end{proof}

>>>>

	\subsection{Tucker's Criterion}\jongmin{notation not yet changed}
	We derive that the PLS components obtained from the first step of the proposed algorithm satisfy the Tucker's Criterion \citep{Tucker1938} extended to our scalar-on-hybrid regression model setting \eqref{eq: Hybrid functional model}. We omit $l$ in the notations and first define the cross-covariance term between the response and the functional predictors (observed versions if $l=1$ and residualized versions if $l \ge 2$): 
	\begin{equation}\label{def:cross_covariance_yx}
\sigma_{YX}
=
 \bigl(
 \sigma_{YX, 1}(t) , \ldots, \sigma_{YX, K}(t) 
 \bigr)
  :=
\bigl(
\mathbb{E}
[Y_1 X_{11}]
,\ldots, 
\mathbb{E}
[Y_1  X_{1K}]
\bigr)   \in \mathcal{F}^{\otimes K},
	\end{equation}
and between the response and the scalar predictor:
\begin{equation*}
\sigma_{YZ} 
= 
(\sigma_{YZ,1}, \ldots, \sigma_{YZ,p})^\top 
:= 
\bigl(
\mathbb{E}[Y_1 Z_{11}]
,\ldots,
\mathbb{E}[Y_1 Z_{1p}]
\bigr)^\top %= E(Y\mathbf{Z})
\in \mathbb{R}^p.
\end{equation*}
Based on these definitions, we define the cross-covariance term between the response and the hybrid predictor as
\begin{equation*}
\Sigma_{YW} :=  
\mathbb{E}
[ Y_1 W_1 ]
=
\bigl(
\sigma_{YX, 1} , \ldots, \sigma_{YX, K} , \, \sigma_{YZ}
\bigr) \in \mathcal{H}.
\end{equation*}
Based on these definitions, we introduce two cross-covariance operators and their properties.
\begin{lemma}\label{lemma:cross_cov_functional}
Define the operator $
\mathcal{C}_{YW}
=
\mathbb{E}(W_1 \otimes_\mathcal{H} Y_1)
: \mathcal{H} \rightarrow \mathbb{R}$ such that for any $h=(f_1, \ldots, f_K, \mathbf{v}) \in \mathcal{H}$, it maps $h$ to a real number as: \begin{equation*} \mathcal{C}_{YW} h := \mathbb{E} \left[ \langle W_1, h \rangle_\mathcal{H} Y_1 \right] = \langle \Sigma_{YW}, h \rangle_\mathcal{H} = \sum \limits_{k=1}^K \int_0^1 \sigma_{YX, k}(t) \, f_k(t) \, dt + \sigma_{YW}^\top \mathbf{v}. \end{equation*} 
\jongmin{instead of showing $\mathcal{U}$ is compact, let's show $\mathcal{C}_{YW}$ is compact}If there exist finite constants $Q_1$ and $Q_2$ such that
\begin{align}\label{condition_for_compact}
	\max \limits_{k=1, \ldots, K} \sup \limits_{t \in [0,1]} \sigma_{YX, k}^{2}(t) < Q_1 \quad \textnormal{and} \quad  \max_{r=1, \ldots, p} \sigma_{YZ, r}^2 < Q_2,
\end{align}
the operator $\mathcal{C}_{YW}$ is a compact operator.
\end{lemma}
Proof of Lemma \ref{lemma:cross_cov_functional}
is provided in Appendix \ref{section:proof:lemma:cross_cov_functional}.
\begin{lemma}\label{lemma:cross_cov_operator}
Define $\mathcal{C}_{WY} = \mathbb{E}[Y \otimes W]: \mathbb{R} \rightarrow \mathcal{H}$,
which	maps  $d \in \mathbb{R}$ to a hybrid object in $\mathcal{H}$ as follows:
	%%%%
	\begin{align*}		
		\mathcal{C}_{WY} d 
		:= 
		\mathbb{E} 
		\bigl[
		\langle Y_1, d \rangle W_1 
		\bigr] 
		= 
		\mathbb{E}
		[ Y_1 W_1 d ] =  d \, \Sigma_{Y W}.
	\end{align*}
Then	$\mathcal{C}_{WY}$ is an adjoint operator of $\mathcal{C}_{YW}$. That is, $\mathcal{C}_{WY} = \mathcal{C}_{YW}^\ast$.
\end{lemma}

The proof of Lemma \ref{lemma:cross_cov_operator} is provided in Appendix \ref{section:proof:lemma:cross_cov_operator}.

Based on these two operators, we define a positive   operator as follows:
\begin{lemma}[Composite cross-covariance operator]\label{lemma:composite_cross_cov}
Define
	$\mathcal{U} := \mathcal{C}_{WY} \circ \mathcal{C}_{YW}: \mathcal{H} \rightarrow \mathcal{H}$ as an operator which performs the following mapping:
	\begin{equation*}
		\mathcal{U} h 
		= 
		\mathcal{C}_{WY} ( \mathcal{C}_{YW} h) 
		=
		\mathcal{C}_{YW} (\langle \Sigma_{YW}, h \rangle_\mathcal{H}) 
		=  
		\Sigma_{YW} \, \langle \Sigma_{YW}, h \rangle_\mathcal{H} 
		= 
		(\Sigma_{YW} \otimes \Sigma_{YW}) h.
	\end{equation*}
		In other words, $\mathcal{U} = \Sigma_{YW} \otimes \Sigma_{YW}$.
		Then $\mathcal{U}$
	 is a self-adjoint, positive-semidefinite, and compact operator.
\end{lemma}
The proof of Lemma \ref{lemma:composite_cross_cov} is provided in Appendix \ref{section:proof:lemma:composite_cross_cov}.

@@

\section{Proof of Lemma \ref{lemma:composite_cross_cov}}\label{section:proof:lemma:composite_cross_cov}
\begin{proof}We show that $\mathcal{U}$ is self-adjoint, positive-semidefinite, and compact, in turn.

\medskip
\noindent
\textit{Self-adjoint.}~
For any $h_1, h_2 \in \mathcal{H}$, we have
\begin{align*}
\langle 
\mathcal{U} h_1, \, h_2 
\rangle_\mathcal{H}  
= 
\bigl \langle 
	\langle h_1, \Sigma_{YW} \rangle_\mathcal{H} \Sigma_{YW},
	\,
	h_2
\bigr \rangle_\mathcal{H}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&= \langle h_1, \Sigma_{YW} \rangle_\mathcal{H} \langle \Sigma_{YW}, h_2 \rangle_\mathcal{H}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\\&= 
\bigl \langle 
h_1,
\,
 \langle \Sigma_{YW}, h_2 \rangle_\mathcal{H} \Sigma_{YW} 
\bigr \rangle_\mathcal{H}
\\ 	&= \langle h_1, \mathcal{U} h_2 \rangle_\mathcal{H}.
\end{align*}

\medskip
\noindent
\textit{Positive-semidefinite.}~
For every $h \in \mathcal{H}$, we have
\begin{equation*}
\langle 
	\mathcal{U} h,
	\,
	h
\rangle_\mathcal{H}
=
\bigl \langle 
	\langle 
		\Sigma_{YW},\, h \rangle_{\mathcal{H}}  \Sigma_{YW}, \, h
\bigr \rangle_\mathcal{H}
	= \langle \Sigma_{YW}, \, h \rangle_\mathcal{H}^2 \ge 0.
\end{equation*}

\medskip
\noindent
\textit{Compact.}~
	By Lemma \ref{lemma:cross_cov_functional}, $\mathcal{C}_{YW}$ is a compact operator.
	By  Theorem 4.1.3 of \citet{hsingTheoreticalFoundationsFunctional2015}, 
	the composition of two operators is compact if either operator is compact. Therefore
	$\mathcal{U} := \mathcal{C}_{WY} \circ \mathcal{C}_{YW}$ is a compact operator.
This completes the proof of Lemma \ref{lemma:composite_cross_cov}.
\end{proof}
